Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Cancer Statistics

### 2020-01-10 paper/paper.Rmd
figure

---
title: "Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Cancer Statistics"
affiliation:
  author-columnar: true         ## one column per author
  #institution-columnar: true  ## one column per institution (multiple autors eventually)
  # wide: true                  ## one column wide author/affiliation fields
  institution:
    - name: Queensland University of Technology
      department: Science and Engineering Faculty
      location: Brisbane, Australia
      email: stephanie.kobakian@qut.edu.au
      mark: 1
      author:
        - name: Stephanie Kobakian
    - name: Monash University
      department: Econometrics and Business Statistics Faculty
      location: Melbourne, Australia
      email: dicook@monash.edu
      mark: 1
      author:
        - name: Dianne Cook
keywords: ["statistics", "visual inference", "geospatial", "population"]
abstract: |
  Choosing a visualisation method may seem to be a simple task. The choropleth map display is commonly used for communicating spatial data. However, when used to communicate the spatial distribution of a statistic the sizes of areas can lead to midinterpretation of the distribution. The visualisation method used to present geospatial data will influence the information derived by map users. Choosing an alternative display will influence the communication of the distribution. The hexagon tile map is presented as an alternative display. Visual inference is used to measure the power of design. The choropleth is used as a comparison point. The hexagon tile map display is an effective visualisation for communicatin=g population related distributions. The hexagon tile map display also allowed identification of the geographic distribution with values monotonically decreasing from North West toward South East areas of Australia.
dev: png
bibliography: mybibfile.bib
output: rticles::ieee_article
editor_options: 
  chunk_output_type: inline
link-citations: yes
---

@@

Figure \ref{detection_compare} displays the average detection rates by type of plot, separately for each trend model. Evaluations on the same data set are connected by a line segment. For both the Three Cities and All Cities trend models performance was substantially better for the hexagon tile map. Viewers detected the data plot substantially more often with the hexagon tile map. One replicate for the All Cities group, had similar detection rates for both plot types, and this is discussed in the supplementary [Lineups] section. Surprisingly, participants could also detect the gradual spatial trend in the NW-SE group from the hexagon tile map. We expected that the choropleth map would be superior for the type of spatial pattern, but the data suggests the hexagon tile map performs slightly better, or equally as well.

@@

tidy(glmer1) %>%
  mutate_at(.vars = c("estimate", "std.error"), round, 2) %>% 
  mutate(p.value = round(p.value, 3)) %>% 
  rowwise() %>% 
  mutate(estimate = case_when(
  p.value <= 0.001 ~ paste0(estimate,"$^{***}$"),
  p.value <= 0.01 ~ paste0(estimate, "$^{** }$"),
  p.value <= 0.05 ~ paste0(estimate, "$^{*  }$"),
  p.value <= 0.01 ~ paste0(estimate, "$^{.  }$"),
  TRUE ~ paste0(estimate))) %>%
  ungroup() %>% 
  filter(!is.na(std.error)) %>% 
  select(Term = term, 
    Estimate = estimate, 
    `Std. Error` = std.error, 
    `P value` = p.value) %>% 
  knitr::kable(format = "latex", escape = FALSE, align= "rcrrr", booktabs = T)
```

The model summary in Table \ref{detect_glmer1} shows that the predicted log odds of the detection rate is 0.46 times that of the choropleth lineups, for the lineups of hexagon tile map. 
The baseline trend model used is the NW-SE trend, the detection rate is expected to decrease comparatively for choropleth displays of the Three Cities and All Cities trends. However, when the interaction is considered the detection rate is expected to increase for the Three Cities and All Cities trends shows in the Hexagon tile map display.

>>>>

---
title: "Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Cancer Statistics"
affiliation:
  author-columnar: true         ## one column per author
  #institution-columnar: true  ## one column per institution (multiple autors eventually)
  # wide: true                  ## one column wide author/affiliation fields
  institution:
    - name: Queensland University of Technology
      department: Science and Engineering Faculty
      location: Brisbane, Australia
      email: stephanie.kobakian@qut.edu.au
      mark: 1
      author:
        - name: Stephanie Kobakian
    - name: Monash University
      department: Econometrics and Business Statistics Faculty
      location: Melbourne, Australia
      email: dicook@monash.edu
      mark: 1
      author:
        - name: Dianne Cook
keywords: ["statistics", "visual inference", "geospatial", "population"]
abstract: |
  Choosing a visualisation method may seem to be a simple task. The choropleth map display is commonly used for communicating spatial data. However, when used to communicate the spatial distribution of a statistic the sizes of areas can lead to midinterpretation of the distribution. The visualisation method used to present geospatial data will influence the information derived by map users. Choosing an alternative display will influence the communication of the distribution. The hexagon tile map is presented as an alternative display. Visual inference is used to measure the power of design. The choropleth is used as a comparison point. The hexagon tile map display is an effective visualisation for communicatin=g population related distributions. The hexagon tile map display also allowed identification of the geographic distribution with values monotonically decreasing from North West toward South East areas of Australia.
dev: png
bibliography: mybibfile.bib
output: rticles::ieee_article
editor_options: 
  chunk_output_type: inline
link-citations: true
---

@@

Figure \ref{fig:detection_compare} displays the average detection rates by type of plot, separately for each trend model. Evaluations on the same data set are connected by a line segment. For both the Three Cities and All Cities trend models performance was substantially better for the hexagon tile map. Viewers detected the data plot substantially more often with the hexagon tile map. One replicate for the All Cities group, had similar detection rates for both plot types, and this is discussed in the supplementary [Lineups] section. Surprisingly, participants could also detect the gradual spatial trend in the NW-SE group from the hexagon tile map. We expected that the choropleth map would be superior for the type of spatial pattern, but the data suggests the hexagon tile map performs slightly better, or equally as well.

@@

glmer_terms <- c("Intercept", "Hex.", "Three Cities", "All Cities", 
  "Hex:Three Cities", "Hex:All Cities")

tidy(glmer1) %>%
  mutate_at(.vars = c("estimate", "std.error"), round, 2) %>% 
  mutate(p.value = round(p.value, 3)) %>% 
  rowwise() %>% 
  mutate(estimate = case_when(
  p.value <= 0.001 ~ paste0(estimate,"$^{***}$"),
  p.value <= 0.01 ~ paste0(estimate, "$^{** }$"),
  p.value <= 0.05 ~ paste0(estimate, "$^{*  }$"),
  p.value <= 0.01 ~ paste0(estimate, "$^{.  }$"),
  TRUE ~ paste0(estimate))) %>%
  ungroup() %>% 
  filter(!is.na(std.error)) %>%
  mutate(term = glmer_terms) %>% 
  select(Term = term, 
    Estimate = estimate, 
    `Std. Error` = std.error, 
    `P value` = p.value) %>% 
  knitr::kable(format = "latex", escape = FALSE, align= "rcrrr", booktabs = T)
```

The model summary in Table \ref{tab:detect_glmer1} shows that the predicted log odds of the detection rate is 0.46 times that of the choropleth lineups, for the lineups of hexagon tile map. 
The baseline trend model used is the NW-SE trend, the detection rate is expected to decrease comparatively for choropleth displays of the Three Cities and All Cities trends. However, when the interaction is considered the detection rate is expected to increase for the Three Cities and All Cities trends shows in the Hexagon tile map display.
